---
title: "BRAF + KRAS flow - Figure S13"
author: "Tina Perica"
date: "2025-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F, warning=F}
library(flowCore)
library(viridisLite)
library(scales)
library(tidyverse)
```

# README
-- Input file for mEGFP-BRAF and mCherry-KRAS
KRAS_BRAF_flow/11.27_all_samples.csv

-- Input file for HA-BRAF and mCherry-KRAS
KRAS_BRAF_flow/11.35_all_samples_single_cells.csv
Note for the HA sample: important for setting transfection gates: must use HA_background for BRAF-HA expression, this is the only one with the correct background, other 2nd only controls do not have this.
For mCherry we can use any non-transfected


```{r}
conditions_in_order  <- c("Full", 
                          "Starved", 
                          "5 min EGF")
```



```{r, message=FALSE, warning=FALSE}
data_mEGFP <- read_csv("KRAS_BRAF_flow/11.27_all_samples.csv") |> 
  add_column("tag" = "mEGFP") |> 
  rename("BRAF" = GFP, "KRAS" = mCherry)
data_HA <- read_csv("KRAS_BRAF_flow/11.35_all_samples_single_cells.csv") |> 
   add_column("tag" = "HA")

data <- rbind(data_mEGFP, data_HA)
```



```{r}
make_logicle_trans <- function(channel, positive_decades = 6) {
  ff <- flowFrame(as.matrix(data %>% select(all_of(channel))))
  # Estimate logicle transform
  trans_list <- estimateLogicle(ff, channels = channel, m = positive_decades)
  logicle_trans <- CytoExploreR_.estimateLogicle(ff, channel, 
                                                 m = positive_decades)
  params <- summary(logicle_trans[[channel]])
  logicle_obj <- logicleTransform(
    t = params$t,
    m = params$m,
    w = params$w,
    a = params$a
  )
  inv_fun <- inverseLogicleTransform(logicle_obj)
  trans_new(
    name = "logicle",
    transform = function(x) logicle_obj@.Data(x),
    inverse = function(x) inv_fun(x),
    domain = c(-1, 1e5),   # adjust based on data
    breaks = extended_breaks(n = 30),
    format = format_format()
  )
}

BRAF_trans <- make_logicle_trans(channel = "BRAF", positive_decades = 6)
KRAS_trans <- make_logicle_trans(channel = "KRAS", positive_decades = 6)
pERK_trans <- make_logicle_trans(channel = "pERK", positive_decades = 6)
custom_breaks_pERK <- c(1e2, 1e3, 1e4, 1e5, 1e6)
custom_breaks_BRAF <- c(1e2, 1e3, 1e4, 1e5)
custom_breaks_KRAS <- c(1e2, 1e3, 1e4, 1e5)
```


```{r}
flow_data_mEGFP <- data |> 
  filter(tag == "mEGFP") |> 
  mutate("DNA_amount" = 
           case_when(
            sample == "BRAFWT" ~ "1200",
            sample == "BRAFAAAA" ~ "1200",
            sample == "BRAFAAAA_KRAS_600" ~"600", 
            sample == "BRAFAAAA_KRAS_750" ~ "750",
            sample == "BRAFWT_KRAS_600" ~ "600",
            sample == "BRAFWT_KRAS_750" ~ "750",
            sample == 'non_transfected' ~ "0",
            .default = "0"
  )) |> 
  mutate(sample =
          case_when(
            sample == "BRAFWT" ~ "mEGFP-BRAF\nWT",
            sample == "BRAFAAAA" ~ "mEGFP-BRAF\nAAAA",
            sample == "BRAFAAAA_KRAS_600" ~ "mEGFP-BRAF\nAAAA +\nmCherry-KRAS",
            sample == "BRAFAAAA_KRAS_750" ~ "mEGFP-BRAF\nAAAA +\nmCherry-KRAS",
            sample == "BRAFWT_KRAS_600" ~ "mEGFP-BRAF\nWT +\nmCherry-KRAS",
            sample == "BRAFWT_KRAS_750" ~ "mEGFP-BRAF\nWT +\nmCherry-KRAS",
            sample == 'non_transfected' ~ "untransfected",
            .default = sample
          )) |>
   mutate("condition" = case_when(
    condition == "full" ~ "Full",
    condition == "starve" ~ "Starved",
    condition == "5min_EGF" ~ "5 min EGF",
    .default = condition
  )) |> 
  filter(! (DNA_amount == "750")) |> 
  select(-DNA_amount)

```


```{r}
flow_data_HA <- data |> 
  filter(tag == "HA") |> 
  mutate(sample =
          case_when(
            sample == "BRAFWT" ~ "HA-BRAF\nWT",
            sample == "BRAFAAAA" ~ "HA-BRAF\nAAAA",
            sample == "BRAFAAAA_KRAS" ~ "HA-BRAF\nAAAA +\nmCherry-KRAS",
            sample == "BRAFWT_KRAS" ~ "HA-BRAF\nWT +\nmCherry-KRAS",
            sample == 'non_transfected' ~ "untransfected",
            .default = sample
          )) |>
   mutate("condition" = case_when(
    condition == "full" ~ "Full",
    condition == "starve" ~ "Starved",
    condition == "5min_EGF" ~ "5 min EGF",
    .default = condition
  ))

```


```{r}
flow_data <- rbind(flow_data_HA, flow_data_mEGFP)
```



```{r}
samples <- c("HA-BRAF\nWT +\nmCherry-KRAS", "HA-BRAF\nAAAA +\nmCherry-KRAS",
             "mEGFP-BRAF\nWT +\nmCherry-KRAS", "mEGFP-BRAF\nAAAA +\nmCherry-KRAS")
flow_data |> 
  filter(sample %in% samples) |> 
  filter(condition %in% conditions_in_order) |> 
  mutate(sample = factor(sample, samples)) |> 
  arrange(sample) |> 
  ggplot(aes(x = BRAF, y = KRAS)) +
  #geom_point(size = 0.05, color = "gray") +
  stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 15, lwd = 0.1) +
  scale_x_continuous(transform = BRAF_trans, 
                    breaks = custom_breaks_BRAF, 
                    name = "mEGFP/HA-BRAF (logicle scale)",
                    labels = function(x) parse(text = paste0("10^", log10(x)))) +
  scale_y_continuous(transform = KRAS_trans, 
                    breaks = custom_breaks_KRAS,
                    name = "mCherry-KRAS (logicle scale)",
                    labels = function(x) parse(text = paste0("10^", log10(x)))) +
  theme_minimal() +
  geom_abline(intercept = 0, slope = 1) +
  facet_grid(~sample) +
  labs(title = "HA/mEGFP-BRAF vs mCherry-KRAS levels in co-transfected cells") +
  theme(
      text = element_text(family = "Helvetica", size = 7),
      axis.title = element_text(size = 7),
      axis.text = element_text(size = 6),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(size = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.1),
      panel.grid.major.x = element_line(linewidth = 0.1),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
    )
ggsave("BRAF-KRAS_coexpression.pdf", width = 6, height = 2.5)
```

# Filter untransfected
```{r}
# BRAF_threshold_mEGFP <- flow_data |> 
#   filter(tag == "mEGFP") |> 
#   filter(sample == "untransfected") |> 
#   arrange(desc(BRAF)) |> 
#   pull(BRAF) |> 
#   (\(.) quantile(., 0.999))()
BRAF_threshold_HA <- flow_data |> 
  filter(tag == "HA") |> 
  filter(condition == "HA_background") |> 
  arrange(desc(BRAF)) |> 
  pull(BRAF) |> 
  (\(.) quantile(., 0.99))()
mCherry_KRAS_threshold <- flow_data |> 
  filter(sample == "untransfected") |> 
  arrange(desc(KRAS)) |> 
  pull(KRAS) |> 
  (\(.) quantile(., 0.999))()

flow_data_exp_mEGFP <- flow_data |>
  filter(tag == 'mEGFP') |>
  filter(BRAF > BRAF_threshold_HA) # use HA for both (higher of the two)

flow_data_exp_HA <- flow_data |>
  filter(tag == 'HA') |>
  filter(BRAF > BRAF_threshold_HA)

flow_data_exp <- rbind(flow_data_exp_HA, flow_data_exp_mEGFP)
```


## Transform
To make the bins, you need to transform the data - otherwise the curve won't match

## Transforming the HA-tag data
```{r, eval = T}
# Choose the channels
channels <- c("BRAF", "KRAS", "pERK")
# Convert data to flowFrame
ff <- flowFrame(as.matrix(flow_data_exp_HA %>% select(all_of(channels))))
# Estimate logicle transform automatically
logicle_trans <- estimateLogicle(ff, channels = channels, m = 6)
# Apply transform to flowFrame
ff_transformed <- transform(ff, logicle_trans)
data_transformed_HA <- as_tibble(exprs(ff_transformed)) |> 
  select("BRAF_logicle" = BRAF, "KRAS_logicle" = KRAS, "pERK_logicle" = pERK)
```


## Transforming the mEGFP-tag data
```{r, eval = T}
# Choose the channels
channels <- c("BRAF", "KRAS", "pERK")
# Convert data to flowFrame
ff <- flowFrame(as.matrix(flow_data_exp_mEGFP %>% select(all_of(channels))))
# Estimate logicle transform automatically
logicle_trans <- estimateLogicle(ff, channels = channels, m = 6)
# Apply transform to flowFrame
ff_transformed <- transform(ff, logicle_trans)
data_transformed_mEGFP <- as_tibble(exprs(ff_transformed)) |> 
  select("BRAF_logicle" = BRAF, "KRAS_logicle" = KRAS, "pERK_logicle" = pERK)
```

Transformed - bind together
```{r, eval = TRUE}
data_transformed <- rbind(data_transformed_HA, data_transformed_mEGFP)
```

```{r}
samples_to_plot <- c("mEGFP-BRAF\nWT", 
                    "mEGFP-BRAF\nAAAA",
                    "mEGFP-BRAF\nWT +\nmCherry-KRAS",
                    "mEGFP-BRAF\nAAAA +\nmCherry-KRAS",
                    "HA-BRAF\nWT", 
                    "HA-BRAF\nAAAA",
                    "HA-BRAF\nWT +\nmCherry-KRAS",
                    "HA-BRAF\nAAAA +\nmCherry-KRAS")
```


# make BRAF bins
```{r}
binned_by_BRAF <- bind_cols(data_transformed, flow_data_exp) |> 
  filter(! sample == "untransfected") |> 
  filter(! sample == "controls") |>
  select(sample, condition, 
         BRAF_logicle, 
         BRAF, KRAS, pERK) |> 
  mutate("BRAF_bins" = cut(BRAF_logicle, breaks = 50)) |> 
  group_by(sample, condition, BRAF_bins) |>
  reframe("median_pERK" = median(pERK),
          "median_BRAF" = median(BRAF),
          "count_points" = n()) |> 
  filter(sample %in% samples_to_plot) |>  
  filter(condition %in% conditions_in_order) |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(sample, condition, BRAF_bins) |> 
  filter(count_points > 20)
```


```{r, eval = T}
colors_for_figure <- c("mEGFP-BRAF\nWT" = "#335b8e", 
            "mEGFP-BRAF\nAAAA" = "#fe6d7e", 
            "mEGFP-BRAF\nAAAA +\nmCherry-KRAS" = "#fe6d7e", 
            "mEGFP-BRAF\nWT +\nmCherry-KRAS" ="#335b8e",
            "HA-BRAF\nWT" = "#335b8e", 
            "HA-BRAF\nAAAA" = "#fe6d7e", 
            "HA-BRAF\nAAAA +\nmCherry-KRAS" = "#fe6d7e", 
            "HA-BRAF\nWT +\nmCherry-KRAS" ="#335b8e"
)
```




# Plot BRAF vs pERK in +/- KRAS samples

```{r}
BRAF_lim <- 12000
pERK_lim <- 9000
flow_data_exp |> 
  filter(sample %in% samples_to_plot) |>  
  filter(condition %in% conditions_in_order) |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(sample, condition) |> 
  ggplot(aes(x = BRAF, y = pERK)) +
  scale_x_continuous(transform = BRAF_trans, breaks = custom_breaks_BRAF,
      name = "HA/mEGFP-BRAF abundance (logicle scale)",
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
  scale_y_continuous(transform = pERK_trans, breaks = custom_breaks_pERK,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)",
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
  geom_point(size = 0.05, alpha = 0.2, color = "gray") +
  stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
  scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
  geom_point(data = binned_by_BRAF,
           aes(x = median_BRAF, y = median_pERK, color = sample), size = 0.3) +
  geom_smooth(data = binned_by_BRAF,
          aes(x = median_BRAF, y = median_pERK, color = sample), 
          span = 0.75, se = FALSE, linewidth = 0.75) +
  scale_color_manual(values = colors_for_figure) +
  facet_grid(sample~condition) +
  geom_hline(yintercept = pERK_lim, linetype = "dashed", lwd = 0.1) +
  geom_vline(xintercept = BRAF_lim, linetype = "dashed", lwd = 0.1) +
  theme_bw() +
  theme(
      text = element_text(family = "Helvetica", size = 6),
      axis.title = element_text(size = 6),
      axis.text = element_text(size = 6),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
  )
ggsave("KRAS_coexp_BRAF_pERK_flow.pdf", width = 4, height = 5.8)
```




# Plot only HA samples, otherwise too repetitive and the figure is too large

```{r}
samples_to_plot <- c(
                    "HA-BRAF\nWT", 
                    "HA-BRAF\nAAAA",
                    "HA-BRAF\nWT +\nmCherry-KRAS",
                    "HA-BRAF\nAAAA +\nmCherry-KRAS")
```


# make BRAF bins
```{r}
binned_by_BRAF <- bind_cols(data_transformed, flow_data_exp) |> 
  filter(! sample == "untransfected") |> 
  filter(! sample == "controls") |>
  select(sample, condition, 
         BRAF_logicle, 
         BRAF, KRAS, pERK) |> 
  mutate("BRAF_bins" = cut(BRAF_logicle, breaks = 50)) |> 
  group_by(sample, condition, BRAF_bins) |>
  reframe("median_pERK" = median(pERK),
          "median_BRAF" = median(BRAF),
          "count_points" = n()) |> 
  filter(sample %in% samples_to_plot) |>  
  filter(condition %in% conditions_in_order) |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(sample, condition, BRAF_bins) |> 
  filter(count_points > 20)
```


```{r, eval = T}
colors_for_figure <- c(
            "HA-BRAF\nWT" = "#335b8e", 
            "HA-BRAF\nAAAA" = "#fe6d7e", 
            "HA-BRAF\nAAAA +\nmCherry-KRAS" = "#fe6d7e", 
            "HA-BRAF\nWT +\nmCherry-KRAS" ="#335b8e"
)
```

# Plot BRAF vs pERK in +/- KRAS samples

```{r}
BRAF_lim <- 12000
pERK_lim <- 9000
flow_data_exp |> 
  filter(sample %in% samples_to_plot) |>  
  filter(condition %in% conditions_in_order) |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(sample, condition) |> 
  ggplot(aes(x = BRAF, y = pERK)) +
  scale_x_continuous(transform = BRAF_trans, breaks = custom_breaks_BRAF,
      name = "HA/mEGFP-BRAF abundance (logicle scale)",
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
  scale_y_continuous(transform = pERK_trans, breaks = custom_breaks_pERK,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)",
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
  geom_point(size = 0.05, alpha = 0.2, color = "gray") +
  stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
  scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
  geom_point(data = binned_by_BRAF,
           aes(x = median_BRAF, y = median_pERK, color = sample), size = 0.3) +
  geom_smooth(data = binned_by_BRAF,
          aes(x = median_BRAF, y = median_pERK, color = sample), 
          span = 0.75, se = FALSE, linewidth = 0.75) +
  scale_color_manual(values = colors_for_figure) +
  facet_grid(sample~condition) +
  geom_hline(yintercept = pERK_lim, linetype = "dashed", lwd = 0.1) +
  geom_vline(xintercept = BRAF_lim, linetype = "dashed", lwd = 0.1) +
  theme_bw() +
  theme(
      text = element_text(family = "Helvetica", size = 6),
      axis.title = element_text(size = 6),
      axis.text = element_text(size = 6),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
  )
ggsave("KRAS_coexp_BRAF_pERK_flow_HA_only.pdf", width = 4, height = 3.2)
```










```{r}
samples_to_plot <- c(
                    "HA-BRAF\nWT +\nmCherry-KRAS",
                    "HA-BRAF\nAAAA +\nmCherry-KRAS")
```


# make mCherry-KRAS bins
```{r, eval = F}
binned_by_KRAS <- bind_cols(flow_data_exp, data_transformed) |> 
  filter(sample %in% samples_to_plot) |> 
  select(sample, condition, 
         KRAS_logicle, 
         BRAF, KRAS, pERK) |> 
  mutate("KRAS_bins" = cut(KRAS_logicle, breaks = 50)) |> 
  group_by(sample, condition, KRAS_bins) |>
  reframe("median_pERK" = median(pERK),
          "median_KRAS" = median(KRAS),
          "count_points" = n()) |> 
  filter(condition %in% conditions_in_order) |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(sample, condition, KRAS_bins) |> 
  filter(count_points > 20)
```


```{r, eval = T}
colors_for_figure <- c(
            "HA-BRAF\nAAAA +\nmCherry-KRAS" = "#fe6d7e", 
            "HA-BRAF\nWT +\nmCherry-KRAS" ="#335b8e"
)
```


# Plot pERK vs mCherry
```{r}
KRAS_lim <- 12000
pERK_lim <- 9000
flow_data_exp |> 
  filter(sample %in% samples_to_plot) |>  
  filter(condition %in% conditions_in_order) |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(sample, condition) |> 
  ggplot(aes(x = KRAS, y = pERK)) +
  scale_x_continuous(transform = KRAS_trans, breaks = custom_breaks_KRAS,
      name = "mCherry-KRAS abundance (logicle scale)",
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
  scale_y_continuous(transform = pERK_trans, breaks = custom_breaks_pERK,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)",
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
  geom_point(size = 0.05, alpha = 0.2, color = "gray") +
  stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
  scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
  geom_point(data = binned_by_KRAS,
           aes(x = median_KRAS, y = median_pERK, color = sample), size = 0.3) +
  geom_smooth(data = binned_by_KRAS,
          aes(x = median_KRAS, y = median_pERK, color = sample), 
          span = 0.75, se = FALSE, linewidth = 0.75) +
  scale_color_manual(values = colors_for_figure) +
  facet_grid(sample~condition) +
  geom_hline(yintercept = pERK_lim, linetype = "dashed", lwd = 0.1) +
  geom_vline(xintercept = KRAS_lim, linetype = "dashed", lwd = 0.1) +
  theme_bw() +
  theme(
      text = element_text(family = "Helvetica", size = 6),
      axis.title = element_text(size = 6),
      axis.text = element_text(size = 6),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
  )
ggsave("KRAS_coexp_KRAS_pERK_flow_HA_samples_only.pdf", width = 4, height = 2)
```






## This is for 3D plot (HA-BRAF vs mCherry-KRAS and pERK in color)

## Transform
To make the bins, you need to transform the data - otherwise the curve won't match
```{r}
flow_data_exp <- flow_data |> 
  filter(KRAS > mCherry_KRAS_threshold) |> 
  filter(BRAF > BRAF_threshold_HA)
# Choose the channels
channels <- c("BRAF", "KRAS", "pERK")
# Convert data to flowFrame
ff <- flowFrame(as.matrix(flow_data_exp %>% select(all_of(channels))))
# Estimate logicle transform automatically
logicle_trans <- estimateLogicle(ff, channels = channels, m = 6)
# Apply transform to flowFrame
ff_transformed <- transform(ff, logicle_trans)
data_transformed <- as_tibble(exprs(ff_transformed)) |> 
  select("BRAF_logicle" = BRAF, "KRAS_logicle" = KRAS, "pERK_logicle" = pERK)
```

```{r}
binned_by_all <- bind_cols(flow_data_exp, data_transformed) |> 
  select(sample, condition, 
         KRAS_logicle, BRAF_logicle, pERK_logicle,
         BRAF, KRAS, pERK) |> 
  mutate("KRAS_bins" = cut(KRAS_logicle, breaks = 15)) |> 
  mutate("BRAF_bins" = cut(BRAF_logicle, breaks = 15)) |> 
  group_by(sample, condition, KRAS_bins, BRAF_bins) |>
  reframe("median_pERK" = median(pERK),
          "median_KRAS" = median(KRAS),
          "median_BRAF" = median(BRAF),
          "count_points" = n()) |> 
  arrange(sample, condition, KRAS_bins, BRAF_bins) |> 
  filter(count_points > 10)
```


```{r}
samples_to_plot <- c(
                    "HA-BRAF\nWT +\nmCherry-KRAS",
                    "HA-BRAF\nAAAA +\nmCherry-KRAS")
```

```{r}
binned_by_all |> 
  filter(sample %in% samples_to_plot) |> 
  filter(condition == "5 min EGF") |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  arrange(sample) |> 
  ggplot(aes(x = median_KRAS, y = median_BRAF, color = median_pERK, size = count_points)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_b() +
  facet_grid(~sample) +
  theme(
    text = element_text(family = "Helvetica", size = 7),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 6),
    axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
    axis.ticks = element_line(linewidth = 0.07),
    axis.ticks.length = unit(0.05, 'cm'),
    legend.position = 'right',
    axis.line = element_line(linewidth = 0.1),
    panel.grid.major.y = element_line(linewidth = 0.15),
    panel.grid.major.x = element_line(linewidth = 0.15),
    panel.border = element_rect(fill = NA, linewidth = 0.2),
    strip.background = element_rect(color = "gray", fill = "white", size = 0.3)
  ) +
  scale_x_continuous(
    trans = KRAS_trans, breaks = custom_breaks_KRAS,
    name = "mCherry-KRAS (Logicle scale)",
    labels = function(x) parse(text = paste0("10^", log10(x)))
  ) +
  scale_y_continuous(
    trans = BRAF_trans, breaks = custom_breaks_BRAF,
    name = "HA-BRAF (Logicle scale)",
    labels = function(x) parse(text = paste0("10^", log10(x)))
  ) 

ggsave("binned_HA-BRAF_and_mCherry-KRAS_pERK_values_logicle.pdf", height = 3, width = 7)
```

```{r}
samples_to_plot <- c(
                    "mEGFP-BRAF\nWT +\nmCherry-KRAS",
                    "mEGFP-BRAF\nAAAA +\nmCherry-KRAS")
```

```{r}
binned_by_all |> 
  filter(sample %in% samples_to_plot) |> 
  filter(condition == "5 min EGF") |> 
  mutate("sample" = factor(sample, samples_to_plot)) |> 
  arrange(sample) |> 
  ggplot(aes(x = median_KRAS, y = median_BRAF, color = median_pERK, size = count_points)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_b() +
  facet_grid(~sample) +
  theme(
    text = element_text(family = "Helvetica", size = 7),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 6),
    axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
    axis.ticks = element_line(linewidth = 0.07),
    axis.ticks.length = unit(0.05, 'cm'),
    legend.position = 'right',
    axis.line = element_line(linewidth = 0.1),
    panel.grid.major.y = element_line(linewidth = 0.15),
    panel.grid.major.x = element_line(linewidth = 0.15),
    panel.border = element_rect(fill = NA, linewidth = 0.2),
    strip.background = element_rect(color = "gray", fill = "white", size = 0.3)
  ) +
  scale_x_continuous(
    trans = KRAS_trans, breaks = custom_breaks_KRAS,
    name = "mCherry-KRAS (Logicle scale)",
    labels = function(x) parse(text = paste0("10^", log10(x)))
  ) +
  scale_y_continuous(
    trans = BRAF_trans, breaks = custom_breaks_BRAF,
    name = "HA-BRAF (Logicle scale)",
    labels = function(x) parse(text = paste0("10^", log10(x)))
  ) 

ggsave("binned_mEGFP-BRAF_and_mCherry-KRAS_pERK_values_logicle.pdf", height = 3, width = 7)
```











```{r}
binned_by_all |> 
  filter(condition == "5 min EGF") |> 
  ggplot(aes(x = median_mCherry, y = median_GFP, color = median_pERK, size = count_points)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_b() +
  facet_grid(~sample) +
  theme(
    text = element_text(family = "Helvetica", size = 7),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 6),
    axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
    axis.ticks = element_line(linewidth = 0.07),
    axis.ticks.length = unit(0.05, 'cm'),
    legend.position = 'right',
    axis.line = element_line(linewidth = 0.1),
    panel.grid.major.y = element_line(linewidth = 0.15),
    panel.grid.major.x = element_line(linewidth = 0.15),
    panel.border = element_rect(fill = NA, linewidth = 0.2),
    strip.background = element_rect(color = "gray", fill = "white", size = 0.3)
  ) +
  scale_x_continuous(
    trans = mCherry_trans, breaks = custom_breaks,
    name = "mCherry-KRAS (Logicle scale)",
    labels = function(x) parse(text = paste0("10^", log10(x)))
  ) +
  scale_y_continuous(
    trans = GFP_trans, breaks = custom_breaks,
    name = "mEGFP-BRAF (Logicle scale)",
    labels = function(x) parse(text = paste0("10^", log10(x)))
  ) 

ggsave("binned_GFP_and_mCherry_pERK_values_EGF_only.pdf", height = 3, width = 7)
```


