---
title: "BRAF signaling flow data analysis: data with HA tag for revised main figure"
author: "Tina Perica"
date: "2025-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(flowCore)
library(scales)
library(tidyverse)
```

# README
Input files for this code are:
## Data with mEGFP-BRAF
--- for pERK
BRAF_flow/11.20_all_samples.csv
BRAF_flow/11.21_all_samples.csv
BRAF_flow/11.22_all_samples.csv
--- for pMEK
BRAF_flow/supplementary_controls/11.25_all_samples_live.csv
## Data with HA-BRAF
BRAF_flow/HA_tag_data/11.28_all_samples.csv
BRAF_flow/HA_tag_data/11.29_all_samples.csv
BRAF_flow/HA_tag_data/11.32_all_samples.csv
BRAF_flow/HA_tag_data/11.33_all_samples.csv

## Define here how you order samples and conditions (important for plots)
```{r}
conditions_in_order  <- c("Full", 
                          "Starved", 
                          "5 min EGF")
```


# Read, merge and tidy up flow data 
This has already been gated and filtered by area and height to remove cell debris and cell clumps in FlowJo
(based on FSC_A, FSC_H, SSC_A, and SSC_H values)

mEGFP samples (annotate the number of replicates)
```{r, message=FALSE, warning=FALSE}
data1 <- read_delim(file = "11.20_all_samples.csv") |> 
  rename("BRAF" = "GFP") |> 
  add_column("replicate" = "mEGFP_rep1") |> 
  add_column("tag" = "mEGFP") |> 
  filter(! condition == "2nd_only")
mEGFP_samples_data1 <- data1 |> 
  pull(sample) |> unique()
data2 <- read_delim(file = "11.21_all_samples.csv") |>
  rename("BRAF" = "GFP") |> 
  add_column("replicate" = "mEGFP_rep1") |>
  add_column("tag" = "mEGFP") |> 
  mutate("replicate" = case_when(
    sample %in% mEGFP_samples_data1 ~ "mEGFP_rep2",
    .default = replicate
  ))
mEGFP_samples_data2 <- data2 |> 
  pull(sample) |> unique()
data3 <- read_delim(file = "11.22_all_samples.csv") |>
  rename("BRAF" = "GFP") |> 
  add_column("replicate" = "mEGFP_rep1") |>
  add_column("tag" = "mEGFPP") |> 
  mutate("replicate" = case_when(
    (sample %in% mEGFP_samples_data1 & (!sample %in% mEGFP_samples_data2)) ~ "mEGFP_rep2",
    (sample %in% mEGFP_samples_data2 & (!sample %in% mEGFP_samples_data1)) ~ "mEGFP_rep2",
    (sample %in% mEGFP_samples_data1 & sample %in% mEGFP_samples_data2) ~ "mEGFP_rep3",
    .default = replicate
  ))
mEGFP_samples_data3 <- data3 |> 
  pull(sample) |> unique()


data2 |> filter(replicate == 'mEGFP_rep2')
data1 |> pull(sample) |> unique()

data2 |> pull(sample) |> unique()
```

HA-tag samples
```{r, message=FALSE}
data4 <- read_delim(file = "HA_tag_data/11.28_all_samples.csv") |> 
  add_column("replicate" = "HA_rep1") |> 
  add_column("tag" = "HA") |> 
  filter(! condition == "2nd_only")
HA_samples_data1 <- data4 |> 
  pull(sample) |> unique()

data5 <- read_delim(file = "HA_tag_data/11.29_all_samples_single_cells.csv") |>
  add_column("replicate" = "HA_rep1") |>
  add_column("tag" = "HA") |>
  mutate("replicate" = case_when(
    sample %in% HA_samples_data1 ~ "HA_rep2",
    .default = replicate
  ))
HA_samples_data2 <- data5 |> 
  pull(sample) |> unique()

data6 <- read_delim(file = "HA_tag_data/11.32_all_samples_single_cells.csv") |>
  add_column("replicate" = "HA_rep1") |>
  add_column("tag" = "HA") |>
  mutate("replicate" = case_when(
    (sample %in% HA_samples_data1 & (!sample %in% HA_samples_data2)) ~ "HA_rep2",
    (sample %in% HA_samples_data2 & (!sample %in% HA_samples_data1)) ~ "HA_rep2",
    (sample %in% HA_samples_data1 & sample %in% HA_samples_data2) ~ "HA_rep3",
    .default = replicate
  ))
HA_samples_data3 <- data6 |> 
  pull(sample) |> unique()

data7 <- read_delim(file = "HA_tag_data/11.33_all_samples_single_cells.csv") |>
  add_column("replicate" = "HA_rep1") |>
  add_column("tag" = "HA") |>
  mutate("sample" = case_when(
    sample == "HA_BRAF_WT" ~ "HA_BRAFWT",
    .default = sample
  )) |> 
  mutate("replicate" = case_when(
    (sample %in% HA_samples_data1 & 
       (!sample %in% HA_samples_data2) & (!sample %in% HA_samples_data3)) ~ "HA_rep2",
    (sample %in% HA_samples_data2 & 
       (!sample %in% HA_samples_data1) & (!sample %in% HA_samples_data3)) ~ "HA_rep2",
    (sample %in% HA_samples_data3 & 
       (!sample %in% HA_samples_data1) & (!sample %in% HA_samples_data2)) ~ "HA_rep2",
    (sample %in% HA_samples_data1 & sample %in% HA_samples_data2 & 
       (!sample %in% HA_samples_data3)) ~ "HA_rep3",
    (sample %in% HA_samples_data1 & sample %in% HA_samples_data3 & 
       (!sample %in% HA_samples_data2)) ~ "HA_rep3",
    (sample %in% HA_samples_data2 & sample %in% HA_samples_data3 & 
       (!sample %in% HA_samples_data1)) ~ "HA_rep3",
    (sample %in% HA_samples_data2 & sample %in% HA_samples_data3 & sample %in% HA_samples_data1) ~ "HA_rep4",
    .default = replicate
  ))
HA_samples_data4 <- data7 |> 
  pull(sample) |> unique()
data <- data1 |> 
  add_row(data2) |> 
  add_row(data3) |> 
  add_row(data4) |> 
  add_row(data5) |> 
  add_row(data6) |> 
  add_row(data7)
```


Sample check
```{r}
data |> 
  select(sample, tag, replicate) |> 
  unique() |> 
  group_by(sample, tag) |> 
  reframe("replicate_count" = n()) |> 
  arrange(tag, sample) 
```




# Flow cytometry data is usually plotted as biexponential transformed data
The 'biexponential' is an over-parameterized inverse of the hyperbolic sine. The function to be inverted takes the form biexp(x) = a*exp(b*(x-w))-c*exp(-d*(x-w))+f with default parameters selected to correspond to the hyperbolic sine.
From a package for flow analysis in R (Bioconductor), flowCore ( https://rdrr.io/bioc/flowCore/man/biexponentialTransform.html )
we can use the logicle
Background: https://docs.flowjo.com/flowjo/graphs-and-gating/gw-transform-overview/gw-transform-benefits/
Logicle paper: PMID: 16604519
Define the function so you can use it to transform the axes on the plot, or transform the data when binning
```{r}
make_logicle_trans <- function(channel = "BRAF", positive_decades = 6) {
  ff <- flowFrame(as.matrix(data %>% select(all_of(channel))))
  # Estimate logicle transform
  trans_list <- estimateLogicle(ff, channels = channel, m = positive_decades)
  logicle_trans <- CytoExploreR_.estimateLogicle(ff, channel, 
                                                 m = positive_decades)
  params <- summary(logicle_trans[[channel]])
  logicle_obj <- logicleTransform(
    t = params$t,
    m = params$m,
    w = params$w,
    a = params$a
  )
  inv_fun <- inverseLogicleTransform(logicle_obj)
  trans_new(
    name = "logicle",
    transform = function(x) logicle_obj@.Data(x),
    inverse = function(x) inv_fun(x),
    domain = c(-1, 1e5),   # adjust based on your data
    breaks = extended_breaks(n = 30),
    format = format_format()
  )
}
BRAF_trans <- make_logicle_trans(channel = "BRAF", positive_decades = 6)
pERK_trans <- make_logicle_trans(channel = "pERK", positive_decades = 6)
custom_breaks <- c(0, 10, 100, 1e3, 1e4, 1e5)
```


```{r}
HA_flow_data <- data |> 
  filter(tag == "HA") |> 
  mutate("sample" = 
          case_when(
            sample == "mEGFP_BRAFWT" ~ "not adjusted mEGFP-BRAF\nWT",
            sample == "mEGFP_BRAFWT_adjusted" ~ "mEGFP-BRAF\nWT",
            sample == "HA_BRAF_481" ~"HA-BRAF\nA481F", 
            sample == "HA_BRAF_481_AAAA" ~ "HA-BRAF\nAAAA-A481F",
            sample == "HA_BRAFWT" ~"HA-BRAF\nWT",
            sample == "HA_BRAF_AAAA" ~"HA-BRAF\nAAAA",
            sample == "HA_BRAF_D565R" ~"HA-BRAF\nD565R", 
            sample == "HA_BRAF_H510F" ~"HA-BRAF\nH510F",
            sample == "HA_BRAF_H570A" ~"HA-BRAF\nH570A",
            sample == "HA_BRAF_H570E" ~"HA-BRAF\nH570E",
            sample == "HA_BRAF_H570R" ~"HA-BRAF\nH570R",
            sample == "HA_BRAF_R506A" ~"HA-BRAF\nR506A",
            sample == "HA_BRAF_R506E" ~"HA-BRAF\nR506E",
            sample == "HA_BRAF_R509H" ~"HA-BRAF\nR509H",
            sample == 'non_transfected' ~ "untransfected",
            .default = sample
          ))
mEGFP_flow_data <- data |> 
  filter(tag == "mEGFP") |> 
  mutate("sample" = 
          case_when(
            sample == "BRAFWT" ~ "mEGFP-BRAF\nWT",
            sample == "BRAFAAAA" ~ "mEGFP-BRAF\nAAAA",
            sample == "BRAF_481A" ~"mEGFP-BRAF\nA481F", 
            sample == "BRAF_481A_AAAA" ~ "mEGFP-BRAF\nAAAA-A481F",
            sample == "BRAF_R509H" ~"mEGFP-BRAF-R509H",
            sample == "BRAF_R509H_AAAA" ~"mEGFP-BRAF-AAAA-R509H",
            sample == "BRAF_D594G" ~"mEGFP-BRAF-D594G", 
            sample == "BRAF_D594G_AAAA" ~"mEGFP-BRAF-AAAA-D594G", 
            sample == "BRAF_K601E" ~"mEGFP-BRAF-K601E", 
            sample == "BRAF_K601E_AAAA" ~"mEGFP-BRAF-AAAA-K601E",
            sample == "BRAF_R509H_K601E" ~"mEGFP-BRAF-R509H-K601E", 
            sample == 'non_transfected' ~ "untransfected",
            sample == 'pMax_GFP' ~ "GFP",
            .default = sample
          )) 
flow_data <- rbind(HA_flow_data, mEGFP_flow_data) |>   
   mutate("condition" = case_when(
    condition == "full" ~ "Full",
    condition == "starve" ~ "Starved",
    condition == "5min_EGF" ~ "5 min EGF",
    .default = condition
  )) |> 
  mutate("uniq_sample" = str_c(tag, "_", sample, "_", replicate)) |> 
  filter(! condition == "transfectionctrl") |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(sample, condition) |> 
  filter(! is.na(condition))  ## this is to remove antibody control samples 
```



# Filter untransfected (for HA and mEGFP separately - different methods to quantify BRAF)
```{r}
transfected_threshold_mEGFP <- flow_data |> 
  filter(tag == "mEGFP") |> 
  filter(sample == "untransfected") |> 
  filter(condition == "Starved") |> 
  arrange(desc(BRAF)) |> 
  pull(BRAF) |> 
  (\(.) quantile(., 0.999))()
flow_data_exp_mEGFP <- flow_data |> 
  filter(tag == "mEGFP") |> 
  filter(BRAF > transfected_threshold_mEGFP)


transfected_threshold_HA <- flow_data |> 
  filter(tag == "HA") |> 
  filter(sample == "untransfected") |> 
  filter(condition == "Starved") |> 
  arrange(desc(BRAF)) |> 
  pull(BRAF) |> 
  (\(.) quantile(., 0.999))()
flow_data_exp_HA <- flow_data |> 
  filter(tag == "HA") |> 
  filter(BRAF > transfected_threshold_HA)

flow_data_exp <- rbind(flow_data_exp_mEGFP, flow_data_exp_HA)
```


# Check which samples and conditions and replicates we have
```{r}
flow_data_exp |> 
  select(sample, condition, replicate, uniq_sample) |> 
  unique() |> 
  arrange(sample, replicate, condition)
```

## Transform
To make the bins, you need to transform the data - otherwise the curve won't match
```{r}
# Choose the channels
channels <- c("BRAF", "pERK")
# Convert data to flowFrame
ff <- flowFrame(as.matrix(flow_data_exp %>% select(all_of(channels))))
# Estimate logicle transform automatically
logicle_trans <- estimateLogicle(ff, channels = channels, m = 6)
# Apply transform to flowFrame
ff_transformed <- transform(ff, logicle_trans)
data_transformed <- as_tibble(exprs(ff_transformed)) |> 
  select("BRAF_logicle" = BRAF, "pERK_logicle" = pERK)
```


# make GFP bins
```{r}
binned_by_BRAF <- bind_cols(data_transformed, flow_data_exp) |> 
  select(sample, replicate, uniq_sample, condition, 
         BRAF_logicle, 
         BRAF, pERK) |> 
  mutate("BRAF_bins" = cut(BRAF_logicle, breaks = 50)) |> 
  group_by(sample, replicate, condition, BRAF_bins) |>
  reframe("median_pERK" = median(pERK),
          "median_BRAF" = median(BRAF),
          "count_points" = n()) |> 
  arrange(sample, replicate, condition, BRAF_bins) |> 
  filter(count_points > 10)
```


## Define the four quadrants
```{r}
total_number_of_cells_per_sample <- flow_data_exp |> 
  group_by(tag, sample, condition, replicate) |> 
  summarise("cell_count" = n())
  
quadrant_percent <- flow_data_exp |> 
  mutate("quadrant" = case_when(
    (BRAF < 5000 & pERK < 7000) ~ "Q3",
    (BRAF > 5000 & pERK < 7000) ~ "Q2",
    (BRAF > 5000 & pERK > 7000) ~ "Q1",
    (BRAF < 5000 & pERK > 7000) ~ "Q4"
  )) |> 
  group_by(tag, sample, condition, replicate, quadrant) |> 
  summarise("quadrant_count" = n()) |> 
  inner_join(x = _, y = total_number_of_cells_per_sample, 
             by = c("tag", "sample", "condition", "replicate")) |> 
  mutate("Q_percent" = round(quadrant_count/cell_count * 100, 2))

write_tsv(quadrant_percent, "quadrant_percent.txt")
```





## Plots for paper (Figure 3 and then the 2 other replicates as supplementary)

```{r}
samples_for_paper <- c("mEGFP-BRAF\nWT", "HA-BRAF\nWT", "HA-BRAF\nAAAA", "HA-BRAF\nA481F", "HA-BRAF\nAAAA-A481F")
flow_data_exp_for_paper <- flow_data_exp |> 
   filter(sample %in% samples_for_paper)

```
# Make flow datasets for each figure

### Main manuscript - Figure 3
```{r}
samples_in_order_Fig3 <- c(
  "mEGFP-BRAF\nWT",
  "HA-BRAF\nWT",
  "HA-BRAF\nAAAA", 
  "HA-BRAF\nA481F", 
  "HA-BRAF\nAAAA-A481F")
colors_Fig3 <- c(
  "HA-BRAF\nWT" = "#335b8e",
  "HA-BRAF\nAAAA" = "#fe6d7e", 
  "HA-BRAF\nA481F" = "#f19167", 
  "HA-BRAF\nAAAA-A481F" ="#e8aa9d", 
  "mEGFP-BRAF\nWT" = "#6ca18f")

flow_data_Fig3 <- flow_data_exp |> 
  filter(tag == "HA") |> 
  filter(sample %in% samples_in_order_Fig3) |> 
  mutate("sample" = factor(sample, samples_in_order_Fig3))
```


#### Samples for Figure SXX - flow data with mEGFP
```{r}
samples_in_order_Fig_SXX <- c(
  "mEGFP-BRAF\nWT", 
  "mEGFP-BRAF\nAAAA", 
  "mEGFP-BRAF\nA481F", 
  "mEGFP-BRAF\nAAAA-A481F", 
  "untransfected",
  "GFP")
```



```{r}
plot_single_replicate <- function(rep_num, 
                                  samples_in_order, colors,
                                  y_axis_label,
                                  plot_prefix) {
  single_replicate <- flow_data_exp |> 
    filter(sample %in% samples_in_order) |> 
    filter(replicate == rep_num) |> 
    mutate("sample" = factor(sample, samples_in_order)) |> 
    arrange(sample, condition)
  binned_by_BRAF_single_replicate <- binned_by_BRAF |> 
    filter(sample %in% samples_in_order) |> 
    filter(replicate == rep_num) |> 
    mutate("sample" = factor(sample, samples_in_order)) |> 
    arrange(sample, condition)
  ranges <- single_replicate |>
    group_by(sample, condition) |>
    summarise(
      xmin = 1,
      xmax = 1000000,
      ymin = 1,
      ymax = 1000000,
      .groups = "drop"
    ) |>
    mutate(
      x_inset = 5,
      y_inset = 5,
      x_left  = 5,
      x_right = 1e5,
      y_bot   = 5,
      y_top   = 10e5
    )

  quad_positions <- tibble(
    quadrant = c("Q1", "Q2", "Q3", "Q4"),
    x_side   = c("right", "left", "left", "right"),
    y_side   = c("top",  "top",   "bot",  "bot")
  )

  quad_labels <- quadrant_percent |>
    filter(replicate == rep_num, sample %in% samples_in_order) |>
    mutate(label = scales::percent(Q_percent/100, accuracy = 0.1)) |>
    left_join(quad_positions, by = "quadrant") |>
    left_join(ranges, by = c("sample","condition")) |>
    mutate(
      x = if_else(x_side == "left",  x_left,  x_right),
      y = if_else(y_side == "top",   y_top,   y_bot),
      hjust = if_else(x_side == "left",  0, 1),
      vjust = if_else(y_side == "top",   1, 0)
    )
  
  single_replicate |> 
    ggplot(aes(x = pERK, y = BRAF)) +
    geom_point(size = 0.05, alpha = 0.2, color = "gray") +
    stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
    scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
    geom_point(data = binned_by_BRAF_single_replicate,
           aes(x = median_pERK, y = median_BRAF, color = sample), size = 0.3) +
    geom_smooth(data = binned_by_BRAF_single_replicate,
          aes(x = median_pERK, y = median_BRAF, color = sample), 
          span = 0.5, linewidth = 0.5, orientation = "y") +
    scale_color_manual(values = colors) +
    facet_grid(sample~condition) +
    theme_bw() +
    theme(
      text = element_text(family = "Helvetica", size = 7),
      axis.title = element_text(size = 7),
      axis.text = element_text(size = 7),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
    ) +
    scale_x_continuous(transform = pERK_trans, breaks = custom_breaks,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)") +
    scale_y_continuous(transform = BRAF_trans, breaks = custom_breaks,
      name = y_axis_label) +
    geom_hline(yintercept = 5000, linetype = "dashed", lwd = 0.1) +
    geom_vline(xintercept = 7000, linetype = "dashed", lwd = 0.1) +
    geom_text(
      data = quad_labels,
      aes(x = x, y = y, label = label, hjust = hjust, vjust = vjust),
      size = 2.2,
      fontface = "bold")
  ggsave(str_c(plot_prefix, rep_num, ".pdf"), device = pdf, height = 9, width = 6)
}
plot_single_replicate(rep_num = "HA_rep1", samples_in_order = samples_in_order_Fig3, colors = colors_Fig3,
                                  y_axis_label = "HA-BRAF (logicle)",
                                  plot_prefix = "pERK_flow_HA_tag_")

plot_single_replicate(rep_num = "HA_rep2", samples_in_order = samples_in_order_Fig3, colors = colors_Fig3,
                                  y_axis_label = "HA-BRAF (logicle)",
                                  plot_prefix = "pERK_flow_HA_tag_")

```


## Plot just the WT (useful for talks/presentations)
```{r}
wt_rep2 <- flow_data_exp_for_paper |> 
  filter(sample == "HA-BRAF\nWT") |> 
  filter(replicate == "rep1")

binned_by_BRAF_wt_rep2 <- binned_by_BRAF |> 
  filter(sample == "HA-BRAF\nWT") |> 
  filter(replicate == "rep1")
wt_rep2 |> 
  ggplot(aes(x = pERK, y = GFP)) +
    geom_point(size = 0.05, alpha = 0.2, color = "gray") +
    stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
    scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
    geom_point(data = binned_by_GFP_wt_rep2,
           aes(x = median_pERK, y = median_GFP, color = sample), size = 0.3) +
    geom_smooth(data = binned_by_GFP_wt_rep2,
          aes(x = median_pERK, y = median_GFP, color = sample), 
          span = 0.5, linewidth = 0.5, orientation = "y") +
    scale_color_manual(values = colors) +
    facet_grid(~condition) +
    theme_bw() +
    theme(
      text = element_text(family = "Helvetica", size = 7),
      axis.title = element_text(size = 7),
      axis.text = element_text(size = 7),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
    ) +
    scale_x_continuous(transform = pERK_trans, breaks = custom_breaks,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)") +
    scale_y_continuous(transform = GFP_trans, breaks = custom_breaks,
      name = "mEGFP-BRAF (logicle)")
ggsave("WT_rep2_flow_for_talks.pdf", height = 2, width = 5)
```

## Plot just the AAAA (useful for talks/presentations)
```{r}
aaaa_rep2 <- flow_data_exp_for_paper |> 
  filter(sample == "mEGFP-BRAF\nAAAA") |> 
  filter(replicate == "rep2")

binned_by_GFP_aaaa_rep2 <- binned_by_GFP |> 
  filter(sample == "mEGFP-BRAF\nAAAA") |> 
  filter(replicate == "rep2")
aaaa_rep2 |> 
  ggplot(aes(x = pERK, y = GFP)) +
    geom_point(size = 0.05, alpha = 0.2, color = "gray") +
    stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
    scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
    geom_point(data = binned_by_GFP_aaaa_rep2,
           aes(x = median_pERK, y = median_GFP, color = sample), size = 0.3) +
    geom_smooth(data = binned_by_GFP_aaaa_rep2,
          aes(x = median_pERK, y = median_GFP, color = sample), 
          span = 0.5, linewidth = 0.5, orientation = "y") +
    scale_color_manual(values = colors) +
    facet_grid(~condition) +
    theme_bw() +
    theme(
      text = element_text(family = "Helvetica", size = 7),
      axis.title = element_text(size = 7),
      axis.text = element_text(size = 7),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
    ) +
    scale_x_continuous(transform = pERK_trans, breaks = custom_breaks,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)") +
    scale_y_continuous(transform = GFP_trans, breaks = custom_breaks,
      name = "mEGFP-BRAF (logicle)")
ggsave("AAAA_rep2_flow_for_talks.pdf", height = 2, width = 5)
```


## Plot just the transfected with GFP only (useful for talks/presentations)
```{r}
gfp_rep1 <- flow_data_exp_for_paper |> 
  filter(sample == "GFP") |> 
  filter(replicate == "rep1")

binned_by_GFP_GFP_rep1 <- binned_by_GFP |> 
  filter(sample == "GFP") |> 
  filter(replicate == "rep1")
gfp_rep1 |> 
  ggplot(aes(x = pERK, y = GFP)) +
  geom_point(size = 0.05, alpha = 0.2, color = "gray") +
    stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
    scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
    geom_point(data = binned_by_GFP_GFP_rep1,
           aes(x = median_pERK, y = median_GFP, color = sample), size = 0.3) +
    geom_smooth(data = binned_by_GFP_GFP_rep1,
          aes(x = median_pERK, y = median_GFP, color = sample), 
          span = 0.5, linewidth = 0.5, orientation = "y") +
    scale_color_manual(values = colors) +
    facet_grid(sample~condition) +
    theme_bw() +
    theme(
      text = element_text(family = "Helvetica", size = 7),
      axis.title = element_text(size = 7),
      axis.text = element_text(size = 7),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
    ) +
    scale_x_continuous(transform = pERK_trans, breaks = custom_breaks,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)") +
    scale_y_continuous(transform = GFP_trans, breaks = custom_breaks,
      name = "mEGFP-BRAF (logicle)")

ggsave("GFP_rep1_flow_for_talks.pdf", height = 2, width = 5)
```


## Plot the pMEK data using the same 2d density plot
```{r, message = FALSE, warning = FALSE}
pmek_data <- read_csv("BRAF_flow/supplementary_controls/11.25_all_samples_live.csv")
pmek_data <- pmek_data |> 
  mutate("sample" = 
          case_when(
            sample == "BRAFWT" ~ "mEGFP-BRAF\nWT",
            sample == "BRAFAAAA" ~ "mEGFP-BRAF\nAAAA",
            sample == "BRAF_481A" ~"mEGFP-BRAF\nA481F", 
            sample == "BRAF_481A_AAAA" ~ "mEGFP-BRAF\nAAAA-A481F",
            sample == 'non_transfected' ~ "untransfected",
            sample == 'pMax_GFP' ~ "GFP",
            .default = sample
          )) |> 
  mutate("condition" = case_when(
    condition == "full" ~ "Full",
    condition == "starve" ~ "Starved",
    condition == "5min_EGF" ~ "5 min EGF",
    .default = condition
  )) |> 
  filter(condition %in% conditions_in_order) |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  mutate("sample" = factor(sample, samples_in_order)) |> 
  arrange(sample, condition) |> 
  filter(! is.na(condition))

pmek_data |> 
  pull(sample) |> unique()
```

# Filter out untransfected cells
```{r}
pmek_GFP_threshold <- pmek_data |> 
  filter(sample == "untransfected") |> 
  arrange(desc(GFP)) |> 
  pull(GFP) |> 
  (\(.) quantile(., 0.999))()
pmek_flow_data_exp <- pmek_data |> 
  filter(GFP > pmek_GFP_threshold) |> 
  filter(!sample == "untransfected") |>
  filter(!sample == "GFP")
```

## Transform function parameters
```{r}
# Choose the channels
channels <- c("GFP", "pMEK")
# Convert data to flowFrame
ff <- flowFrame(as.matrix(pmek_flow_data_exp %>% select(all_of(channels))))
# Estimate logicle transform automatically
logicle_trans <- estimateLogicle(ff, channels = channels, m = 6)
# Apply transform to flowFrame
ff_transformed <- transform(ff, logicle_trans)
mek_data_transformed <- as_tibble(exprs(ff_transformed)) |> 
  select("GFP_logicle" = GFP, "pMEK_logicle" = pMEK)
```


# make GFP bins
```{r}
pmek_binned_by_GFP <- bind_cols(mek_data_transformed, pmek_flow_data_exp) |> 
  select(sample, condition, 
         GFP_logicle, 
         GFP, pMEK) |> 
  mutate("GFP_bins" = cut(GFP_logicle, breaks = 50)) |> 
  group_by(sample, condition, GFP_bins) |>
  reframe("median_pMEK" = median(pMEK),
          "median_GFP" = median(GFP),
          "count_points" = n()) |> 
  arrange(sample, condition, GFP_bins) |> 
  filter(count_points > 10)
```

```{r}
pmek_flow_data_exp |> 
  ggplot(aes(x = pMEK, y = GFP)) +
  geom_point(size = 0.05, alpha = 0.2, color = "gray") +
  stat_density_2d(geom = "polygon", contour = TRUE,
          aes(fill = after_stat(level)), colour = "gray",
          bins = 20, lwd = 0.1) +
  scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
  geom_point(data = pmek_binned_by_GFP,
          aes(x = median_pMEK, y = median_GFP, color = sample), size = 0.3) +
  geom_smooth(data = pmek_binned_by_GFP,
          aes(x = median_pMEK, y = median_GFP, color = sample), 
          span = 0.5, linewidth = 0.5, orientation = "y") +
  scale_color_manual(values = colors) +
  facet_grid(sample~condition) +
  theme_bw() +
  theme(
    text = element_text(family = "Helvetica", size = 7),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 7),
    axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
    axis.ticks = element_line(linewidth = 0.07),
    axis.ticks.length = unit(0.05, 'cm'),
    legend.position = 'none',
    axis.line = element_line(linewidth = 0.1),
    panel.grid.major.y = element_line(linewidth = 0.15),
    panel.grid.major.x = element_line(linewidth = 0.15),
    panel.border = element_rect(fill = NA, linewidth = 0.2),
    strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
  ) +
  scale_x_continuous(transform = pERK_trans, breaks = custom_breaks,
    name = "MAPK/ERK pathway activity - pMEK levels (logicle scale)") +
  scale_y_continuous(transform = GFP_trans, breaks = custom_breaks,
    name = "mEGFP-BRAF (logicle)")

ggsave("pMEK_2d_density_plots.pdf", height = 5, width = 6)
```

