---
title: "Landing pad BRAF flow data figure"
author: "Tina Perica"
date: "2026-03-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(flowCore)
library(scales)
library(tidyverse)
```

# README
Input files for this code are:
## Data with mEGFP-BRAF expressed from landing pad
--- for pERK
Landing_pad_data/11.30_all_samples_single_cells.csv
Landing_pad_data/11.31_all_samples_single_cells.csv


## Define here how you order samples and conditions (important for plots)
```{r}
conditions_in_order  <- c("Full", 
                          "Starved", 
                          "5 min EGF")
```


# Read, merge and tidy up flow data 
This has already been gated and filtered by area and height to remove cell debris and cell clumps in FlowJo
(based on FSC_A, FSC_H, SSC_A, and SSC_H values)

mEGFP samples (annotate the number of replicates)
```{r, message=FALSE, warning=FALSE}
data1 <- read_delim(file = "Landing_pad_data/11.30_all_samples_single_cells.csv") |> 
  add_column("replicate" = "mEGFP_rep1") |> 
  add_column("tag" = "mEGFP") |> 
  filter(! condition == "2nd_only")
mEGFP_samples_data1 <- data1 |> 
  pull(sample) |> unique()
data2 <- read_delim(file = "Landing_pad_data/11.31_all_samples_single_cells.csv") |>
  add_column("replicate" = "mEGFP_rep1") |>
  add_column("tag" = "mEGFP") |> 
  mutate("replicate" = case_when(
    sample %in% mEGFP_samples_data1 ~ "mEGFP_rep2",
    .default = replicate
  ))

data <- data1 |> 
  add_row(data2)
```



Sample check
```{r}
data |> 
  select(sample, tag, replicate) |> 
  unique() |> 
  group_by(sample, tag) |> 
  reframe("replicate_count" = n()) |> 
  arrange(tag, sample) 
```




# Flow cytometry data is usually plotted as biexponential transformed data
The 'biexponential' is an over-parameterized inverse of the hyperbolic sine. The function to be inverted takes the form biexp(x) = a*exp(b*(x-w))-c*exp(-d*(x-w))+f with default parameters selected to correspond to the hyperbolic sine.
From a package for flow analysis in R (Bioconductor), flowCore ( https://rdrr.io/bioc/flowCore/man/biexponentialTransform.html )
we can use the logicle
Background: https://docs.flowjo.com/flowjo/graphs-and-gating/gw-transform-overview/gw-transform-benefits/
Logicle paper: PMID: 16604519
Define the function so you can use it to transform the axes on the plot, or transform the data when binning
```{r}
make_logicle_trans <- function(channel = "BRAF", positive_decades = 6) {
  ff <- flowFrame(as.matrix(data %>% select(all_of(channel))))
  # Estimate logicle transform
  trans_list <- estimateLogicle(ff, channels = channel, m = positive_decades)
  logicle_trans <- CytoExploreR_.estimateLogicle(ff, channel, 
                                                 m = positive_decades)
  params <- summary(logicle_trans[[channel]])
  logicle_obj <- logicleTransform(
    t = params$t,
    m = params$m,
    w = params$w,
    a = params$a
  )
  inv_fun <- inverseLogicleTransform(logicle_obj)
  trans_new(
    name = "logicle",
    transform = function(x) logicle_obj@.Data(x),
    inverse = function(x) inv_fun(x),
    domain = c(-1, 1e5),   # adjust based on your data
    breaks = extended_breaks(n = 30),
    format = format_format()
  )
}
BRAF_trans <- make_logicle_trans(channel = "BRAF", positive_decades = 6)
pERK_trans <- make_logicle_trans(channel = "pERK", positive_decades = 6)
custom_breaks_pERK <- c(1e2, 1e3, 1e4, 1e5, 1e6)
custom_breaks_BRAF <- c(1e2, 1e3, 1e4, 1e5, 1e6)

```


```{r}
flow_data <- data |> 
  mutate("sample" = 
          case_when(
            sample == "BRAFWT_LP" ~ "integrated\nmEGFP-BRAF-WT",
            sample == "BRAFWT_transfected" ~ "transfected\nmEGFP-BRAF-WT",
            sample == "BRAFAAAA_LP" ~"integrated\nmEGFP-BRAF-AAAA", 
            sample == "BRAFAAAA_transfected" ~ "transfected\nmEGFP-BRAF-AAAA",
            sample == 'non_transfected' ~ "untransfected",
            .default = sample
          )) |> 
   mutate("condition" = case_when(
    condition == "full" ~ "Full",
    condition == "starve" ~ "Starved",
    condition == "5min_EGF" ~ "5 min EGF",
    .default = condition
  )) |> 
  mutate("uniq_sample" = str_c(tag, "_", sample, "_", replicate)) |> 
  filter(! condition == "transfectionctrl") |> 
  mutate("condition" = factor(condition, conditions_in_order)) |> 
  arrange(condition) |> 
  filter(! is.na(condition))  ## this is to remove antibody control samples 
```



# Filter untransfected
```{r}
transfected_threshold_mEGFP <- flow_data |> 
  filter(sample == "untransfected") |> 
  filter(condition == "Starved") |> 
  arrange(desc(BRAF)) |> 
  pull(BRAF) |> 
  (\(.) quantile(., 0.999))()
flow_data_exp <- flow_data |> 
  filter(BRAF > transfected_threshold_mEGFP)

```


## Transform
To make the bins, you need to transform the data - otherwise the curve won't match

## Transforming the mEGFP-tag data
```{r, eval = T}
# Choose the channels
channels <- c("BRAF", "pERK")
# Convert data to flowFrame
ff <- flowFrame(as.matrix(flow_data_exp %>% select(all_of(channels))))
# Estimate logicle transform automatically
logicle_trans <- estimateLogicle(ff, channels = channels, m = 6)
# Apply transform to flowFrame
ff_transformed <- transform(ff, logicle_trans)
data_transformed <- as_tibble(exprs(ff_transformed)) |> 
  select("BRAF_logicle" = BRAF, "pERK_logicle" = pERK)
```



# make GFP bins
```{r}
binned_by_BRAF <- bind_cols(data_transformed, flow_data_exp) |> 
  select(tag, sample, replicate, uniq_sample, condition, 
         BRAF_logicle, 
         BRAF, pERK) |> 
  group_by(tag) |> 
  mutate("BRAF_bins" = cut(BRAF_logicle, breaks = 80)) |> 
  group_by(tag, sample, replicate, condition, BRAF_bins) |>
  reframe("median_pERK" = median(pERK),
          "median_BRAF" = median(BRAF),
          "count_points" = n()) |> 
  arrange(sample, replicate, condition, BRAF_bins) |> 
  filter(count_points > 50)


binned_by_BRAF |> 
  group_by(tag) |> 
  arrange(tag, BRAF_bins) |> 
  select(BRAF_bins) |> 
  unique()

binned_by_BRAF |> 
  pull(sample) |> unique()
```



## Define the four quadrants
```{r}
total_number_of_cells_per_sample <- flow_data_exp |> 
  group_by(tag, sample, condition, replicate) |> 
  summarise("cell_count" = n())
```

## Make a plotting function
```{r}
quadrant_function <- function(BRAF_lim = 12000, pERK_lim = 9000, flow_data_for_quadrants) {
  quadrant_percent <- flow_data_for_quadrants |> 
  mutate("quadrant" = case_when(
    (BRAF > BRAF_lim & pERK > pERK_lim) ~ "Q1",
    (BRAF < BRAF_lim & pERK > pERK_lim) ~ "Q2",
    (BRAF < BRAF_lim & pERK < pERK_lim) ~ "Q3",
    (BRAF > BRAF_lim & pERK < pERK_lim) ~ "Q4"
  )) |> 
  group_by(tag, sample, condition, replicate, quadrant) |> 
  summarise("quadrant_count" = n()) |> 
  inner_join(x = _, y = total_number_of_cells_per_sample, 
             by = c("tag", "sample", "condition", "replicate")) |> 
  mutate("Q_percent" = round(quadrant_count/cell_count * 100, 2))
  return(quadrant_percent)
}

make_plotting_dataset <- function(flow_data, binned_data, samples_in_order, rep_num, BRAF_lim, pERK_lim) {
  single_replicate <- flow_data |> 
    filter(sample %in% samples_in_order) |> 
    filter(replicate == rep_num) |> 
    mutate("sample" = factor(sample, samples_in_order)) |> 
    arrange(sample, condition)
  binned_by_BRAF_single_replicate <- binned_data |> 
    filter(sample %in% samples_in_order) |> 
    filter(replicate == rep_num) |> 
    mutate("sample" = factor(sample, samples_in_order)) |> 
    arrange(sample, condition)
  
  # Position labels in logicle-transformed space, then invert to data space.
  # Using larger inset fractions so labels sit near panel corners, not near gate lines.
  pERK_trans_fn <- pERK_trans$transform
  BRAF_trans_fn <- BRAF_trans$transform
  pERK_inv_fn   <- pERK_trans$inverse
  BRAF_inv_fn   <- BRAF_trans$inverse

  max_pERK <- max(flow_data$pERK) + 0.1*max(flow_data$pERK)
  max_BRAF <- max(flow_data$BRAF) + 0.1*max(flow_data$BRAF)
  min_pERK <- min(flow_data$pERK) - 0.1*min(flow_data$pERK)
  min_BRAF <- min(flow_data$BRAF) - 0.1*min(flow_data$BRAF)
  
  x_range_t <- range(BRAF_trans_fn(c(min_BRAF, max_BRAF)), na.rm = TRUE)
  y_range_t <- range(pERK_trans_fn(c(min_pERK, max_pERK)), na.rm = TRUE)
  
  x_left  <- BRAF_inv_fn(x_range_t[1] + 0.01 * diff(x_range_t))
  x_right <- BRAF_inv_fn(x_range_t[2] - 0.01 * diff(x_range_t))
  y_bot   <- pERK_inv_fn(y_range_t[1] + 0.05 * diff(y_range_t))
  y_top   <- pERK_inv_fn(y_range_t[2] - 0.01 * diff(y_range_t))
  

  ranges <- single_replicate |>
    group_by(sample, condition) |>
    summarise(.groups = "drop") |>
    mutate(
      x_left  = x_left,
      x_right = x_right,
      y_bot   = y_bot,
      y_top   = y_top
    )
  
  quad_positions <- tibble(
    quadrant = c("Q1", "Q2", "Q3", "Q4"),
    x_side   = c("right", "left", "left", "right"),
    y_side   = c("top",  "top",   "bot",  "bot")
  )
  quadrant_percent <- quadrant_function(BRAF_lim = BRAF_lim, pERK_lim = pERK_lim, 
                                        flow_data_for_quadrants = single_replicate)
  quad_labels <- quadrant_percent |>
    filter(replicate == rep_num, sample %in% samples_in_order) |>
    mutate("sample" = factor(sample, samples_in_order)) |> 
    mutate(label = scales::percent(Q_percent/100, accuracy = 0.1)) |>
    left_join(quad_positions, by = "quadrant") |>
    left_join(ranges, by = c("sample","condition")) |>
    mutate(
      x = if_else(x_side == "left",  x_left,  x_right),
      y = if_else(y_side == "top",   y_top,   y_bot),
      hjust = if_else(x_side == "left",  0, 1),
      vjust = if_else(y_side == "top",   1, 0)
    ) |> 
  arrange(sample, condition)
  to_plot <- list(
    "single_replicate" = single_replicate,
    "binned_by_BRAF_single_replicate" = binned_by_BRAF_single_replicate,
    "quad_labels" = quad_labels
  )
  return(to_plot)
}
plot_single_replicate <- function(flow_data,
                                  binned_data,
                                  samples_in_order, 
                                  rep_num, 
                                  BRAF_lim, 
                                  pERK_lim,
                                  colors,
                                  x_axis_label,
                                  plot_prefix,
                                  plot_height = 9,
                                  plot_width = 6) {
  
  data_to_plot <- make_plotting_dataset(flow_data, binned_data, samples_in_order, 
                                        rep_num, BRAF_lim = BRAF_lim, pERK_lim = pERK_lim)
  data_to_plot[["single_replicate"]] |> 
    ggplot(aes(x = BRAF, y = pERK)) +
    scale_x_continuous(transform = BRAF_trans, breaks = custom_breaks_BRAF,
      name = x_axis_label,
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
    scale_y_continuous(transform = pERK_trans, breaks = custom_breaks_pERK,
      name = "MAPK/ERK pathway activity - pERK levels (logicle scale)",
      labels = function(x) parse(text = paste0("10^", log10(x)))) +
    geom_point(size = 0.05, alpha = 0.2, color = "gray") +
    stat_density_2d(geom = "polygon", contour = TRUE,
                aes(fill = after_stat(level)), colour = "gray",
                bins = 20, lwd = 0.1) +
    scale_fill_distiller(palette = "Blues", direction = 1, name = "point density") +
    # geom_point(data = data_to_plot[["binned_by_BRAF_single_replicate"]],
    #        aes(x = median_BRAF, y = median_pERK, color = sample), size = 0.3) +
    # geom_smooth(data = data_to_plot[["binned_by_BRAF_single_replicate"]],
    #       aes(x = median_BRAF, y = median_pERK, color = sample), 
    #       span = 0.5, se = FALSE, linewidth = 0.75) +
    scale_color_manual(values = colors) +
    facet_grid(sample~condition) +
    geom_hline(yintercept = pERK_lim, linetype = "dashed", lwd = 0.1) +
    geom_vline(xintercept = BRAF_lim, linetype = "dashed", lwd = 0.1) +
    geom_text(
      data = data_to_plot[["quad_labels"]],
      aes(x = x, y = y, label = label),
      hjust = data_to_plot[["quad_labels"]]$hjust,
      vjust = data_to_plot[["quad_labels"]]$vjust,
      size = 2.2,
      fontface = "bold") +
    theme_bw() +
    theme(
      text = element_text(family = "Helvetica", size = 7),
      axis.title = element_text(size = 7),
      axis.text = element_text(size = 7),
      axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
      axis.ticks = element_line(linewidth = 0.07),
      axis.ticks.length = unit(0.05, 'cm'),
      legend.position = 'none',
      axis.line = element_line(linewidth = 0.1),
      panel.grid.major.y = element_line(linewidth = 0.15),
      panel.grid.major.x = element_line(linewidth = 0.15),
      panel.border = element_rect(fill = NA, linewidth = 0.2),
      strip.background = element_rect(color = "gray", fill = "white", linewidth = 0.3)
    )
  ggsave(str_c(plot_prefix, rep_num, ".pdf"), device = pdf, height = plot_height, width = plot_width)
}
```



## make plots for each figure
1. Make flow datasets for each figure
2. Plot replicate by replicate

### Supplementary Figure X - BRAF-AAAA from a landing pad
```{r}
samples_in_order_FigSX <- c(
  "integrated\nmEGFP-BRAF-WT",
  "transfected\nmEGFP-BRAF-WT",
  "integrated\nmEGFP-BRAF-AAAA", 
  "transfected\nmEGFP-BRAF-AAAA")
colors_FigSX <- c(
  "integrated\nmEGFP-BRAF-WT" = "#335b8e",
  "transfected\nmEGFP-BRAF-WT" = "#335b8e",
  "integrated\nmEGFP-BRAF-AAAA" = "#fe6d7e", 
  "transfected\nmEGFP-BRAF-AAAA" = "#fe6d7e")

plot_single_replicate(flow_data = flow_data_exp, binned_data = binned_by_BRAF,
                      samples_in_order = samples_in_order_FigSX, rep_num = "mEGFP_rep1", 
                      BRAF_lim = 5000, pERK_lim = 9000,
                      colors = colors_FigSX,
                      x_axis_label = "mEGFP-BRAF abundance (logicle scale)", 
                      plot_prefix = "For_rebuttal_landing_pad_flow_",
                      plot_height = 5, plot_width = 5)

plot_single_replicate(flow_data = flow_data_exp, binned_data = binned_by_BRAF,
                      samples_in_order = samples_in_order_FigSX, rep_num = "mEGFP_rep2", 
                      BRAF_lim = 5000, pERK_lim = 9000,
                      colors = colors_FigSX,
                      x_axis_label = "mEGFP-BRAF abundance (logicle scale)", 
                      plot_prefix = "landing_pad_flow_",
                      plot_height = 5, plot_width = 5)

```





